---
title: "How has freedom of movement with the EU affected non-EU migration to the UK?"
author: "Calum Davey"
date: "31/05/2019"
output: md_document
---

```{r, echo=FALSE}
# Read in the data 
d <- read.csv('net-migration-by-citizen.csv')  

# Add EU migration after 2003
m91_03 <- median(d[1:13,5]) # but subtract the median migration before 2003
d$nonEU.plus <- d$Non.EU + d$EU - m91_03
d[1:14,6] <- d[1:14,4]

# Graph of migration to the UK, EU and non-EU 
par(lwd=2)
plot(d[,1],d[,4],typ='l',bty='n',xlab='Year',ylab='Thousands',col='green',ylim=c(0,360))
lines(d[,1],d[,5],col='blue')
abline(v=2004,col='gray50')
p <- recordPlot()
p
legend(x=1991,y=350,legend=c('Free movement','Conventional routes'), fill=c('blue','green'),bty='n',border='white')
```


```{r, echo=FALSE}
p
lines(d[14:nrow(d),1],d[14:nrow(d),6], col='red')
legend(x=1991,y=350,legend=c('Free movement','Conventional routes', 'Missing conventional migrants'), fill=c('blue','green','red'),bty='n',border='white')
```

The total missing non-EU migrants is the sum of the difference between the red and green lines, which is `r format(sum(d[14:nrow(d),6]-d[14:nrow(d),5])*1000,big.mark=',')`. Since the Migration Observatory [estimates that the non-EU foreign-born in the UK in 2017 was only about twixe that figure at 5,677,000](https://migrationobservatory.ox.ac.uk/resources/briefings/migrants-in-the-uk-an-overview/), the 'missing' are `r round(100*sum(d[14:nrow(d),6]-d[14:nrow(d),5])/sum(sum(d[14:nrow(d),6]-d[14:nrow(d),5]), 5677),0)`% of the total non-EU nationals that would be living in the UK without the expansion of access in 2004. Another way of putting it, would be to say that the non-EU born population of the UK in 2017 would have been `r round(100*sum(d[14:nrow(d),6]-d[14:nrow(d),5])/5677,0)`% higher than it was. 

