---
title: "Missing nonEU migration"
author: "Calum Davey"
date: "31/05/2019"
output: html_document
---

Recent [reports from the Office of National Statistics (ONS)](https://www.ons.gov.uk/peoplepopulationandcommunity/populationandmigration/internationalmigration/bulletins/migrationstatisticsquarterlyreport/may2019) have shown that EU migration to the UK has fallen, and that non-EU migrants are taking up the slack. This is interesting for lots of reasons, including implying that [UK employers are willing to make the extra effort to sponsor migrants](https://www.newstatesman.com/politics/staggers/2019/05/when-it-comes-immigration-uk-already-post-brexit-era) when less beurocratic options go down. 

It also implies that over the short term there is roughly a total number of jobs or other social positions that need to be filled by migration each year. This can increase because of the effects of migration on the economy -- I'm not saying there's a limit to the size of the country/economy -- but that takes time. 

I was interested to look at how migration from the EU and non-EU had changed over time, and in particular the effect that the [expansion of the Union from 2004](https://eu.boell.org/en/2014/06/10/europe-after-eastern-enlargement-european-union-2004-2014) had on the number of non-EU migrants to the UK. 

The ONS data comes in multiple tables with lots of detail, but fortunately the [Migration Observatory at Oxford University](https://migrationobservatory.ox.ac.uk/resources/briefings/long-term-international-migration-flows-to-and-from-the-uk/) have compiled the annual net migration data by EU/non-EU, which can be downloaded as a [.csv file](net-migration-by-citizen.csv). 

Having loaded the data into [R](https://www.r-project.org), a free statistical software, I did a couple of small calculations before making the graph (which you can see in the code at the end). In order to get a rough estimate of the number of non-EU migrants that there _would have been_ had the EU not expanded in 2004 (the 'missing' non-EU migrants), I added the number of EU migrants to the number of non-EU migrants after 2004. Since there was net migration from the EU before 2004, and I couldn't work out which came through conventional channels and which were from pre-2004 EU countries, I subtracted the median migration between 1991 and 2003 from each year after 2003 to account for that. If a large proportion of those were from pre-2004 EU countries then the number of 'missing' non-EU migrants will have been underestimated.  

The result is a very simple graph showing how EU and non-EU net migration from 1991 to 2017 (updates to 2019 would be possible but require downloading the ONS data). We can see the strong upward trend in non-EU migration from 1991 to 2004, and subsequent leveling-off and slight decline. In contrast, the blue EU line is relatively flat before 2004, after which it jumps up. 

The red line on the graph shows the 'missing' non-EU migrants; as expected this tracks the EU and non-EU migration. 

```{r, echo=FALSE}
# Read in the data 
d <- read.csv('net-migration-by-citizen.csv')  

# Add EU migration after 2003
m91_03 <- median(d[1:13,5]) # but subtract the median migration before 2003
d$nonEU.plus <- d$Non.EU + d$EU - m91_03
d[1:14,6] <- d[1:14,4]

# Graph of migration to the UK, EU and non-EU 
par(lwd=2)
plot(d[,1],d[,4],typ='l',bty='n',xlab='Year',ylab='Thousands',col='green',ylim=c(0,360))
lines(d[,1],d[,5],col='blue')
abline(v=2004,col='gray50')
lines(d[14:nrow(d),1],d[14:nrow(d),6], col='red')
legend(x=1991,y=350,legend=c('EU','non-EU','Missing-non-EU'), fill=c('blue','green','red'),bty='n',border='white')
```

The total missing non-EU migrants is the sum of the difference between the red and green lines, which is `r format(sum(d[14:nrow(d),6]-d[14:nrow(d),5])*1000,big.mark=',')`. 

This is a crude analysis, and a very rough estimate of the counterfactual effects on non-EU migration had free movement of people not been expanded in 2004. As I said, subtracting the pre-2004 EU migration figure could underestimate the figure, but this number is not high. 

```{r, echo=TRUE, eval=FALSE}
# Read in the data 
  d <- read.csv('net-migration-by-citizen.csv')  

# Add EU migration after 2003
  m91_03 <- median(d[1:13,5]) # but subtract the median migration before 2003
  d$nonEU.plus <- d$Non.EU + d$EU - m91_03
  d[1:14,6] <- d[1:14,4]

# Graph of migration to the UK, EU and non-EU 
  par(lwd=2)
  plot(d[,1],d[,4],typ='l',bty='n',xlab='Year',ylab='Thousands',col='green',ylim=c(0,360))
    lines(d[,1],d[,5],col='blue')
    abline(v=2004,col='gray50')
    lines(d[14:nrow(d),1],d[14:nrow(d),6], col='red')
    legend(x=1991,y=350,legend=c('EU','non-EU','Missing-non-EU'), fill=c('blue','green','red'),bty='n',border='white')
```
