---
title: "PENDA Deworm3 Precision Simulation"
author: "Calum Davey"
date: "04/04/2019"
output: 
  html_document:
    code_folding: hide
---
## Set the design parameters for the study 

The first stage is to set the number of disabled and non-disabled children that will be recruited overall, and the number of clusters. The number of children per cluster is assumed to be equal, although this could be relaxed. 

```{r}
n <- 2000 # 2000 disabled children, 2000 non-disabled 
d <- matrix(,nrow=n*2, ncol=4) # Create data 
d[,1] <- rep(c(0,1),times=c(n,n)) # Code disabled, and non-disabled (1,0)
c <- 40 # total number of clusters 
d[,2] <- rep(rep(seq(1,c,1), times=rep(n/c, times=c)),times=2)
```

## Set assummed population-level parameters 

Next, we set the population-level parameters, e.g. the 'true' paramter values, for each cluster. In this case, that means the true proportion attending school for disabled and non-disabled children. 

```{r}
e0_p <- c(.7) # enrolment of non-disabled children 
e1_p <- c(.5) # enrolment of disabled children 
k <- c(.01) # Set intercluster coefficient of variation
var_b1 <- k/e1_p
var_b2 <- k/e0_p
# Fix true cluster-level means of the outcome 
e1_c <- rnorm(40, e1_p, var_b1^.5)
e0_c <- rnorm(40, e0_p, var_b2^.5)
d[,3] <- c(rep(e0_c, times=rep(n/c, times=c)),
           rep(e1_c, times=rep(n/c, times=c)))
  # Do need a way to avoid values >1 or <0
```

## Generate enrolment outcome 

Finally, we take 1000 samples from the population, essentially simulating doing the planned study 1000 times. 

```{r}
i <- 1
while (i<=1000){
  d <- cbind(d, apply(d, 1, function(x) sample(c(0,1),size=1,prob=c(1-x[3],x[3]))))
  i <- i+1
}
```

## Conduct analysis on simulated data 

Using the 1000 simulations, we can conduct whatever analysis we want and see how the estiamtes vary from one simulated sample to the next. For example, we can estimate the proportion attending and the difference between the proportions for disabled and non disabled children. 

```{r}
# Calculate sample enrolment figures 
p0 <- apply(d[d[,1]==0, 5:ncol(d)], 2, mean)
p1 <- apply(d[d[,1]==1, 5:ncol(d)], 2, mean)
diff <- apply(cbind(p0,p1),1,function(x) x[1]-x[2])
```

We can visualise the level of precision by plotting the results from all 1000 samples. Here we have the proportion in school for children with disabilities, children without disabilities (controls), and the difference between these for each run of the simulation. 

```{r}
hist(p0, xlim=c(0.1,.8),ylim=c(0,50), main='',
     xlab='R or RD', freq=F)
hist(p1, add =T, freq=F)
hist(diff, xlab = 'RD',add=T, freq=F)
lines(density(diff),col='red')
text(x=c(mean(diff),mean(p1),mean(p0)), 
     y=c(50),
     labels=c('Prop.-diff.', 'Prop. CWD',
              'Prop. controls'))
```





