---
title: "PENDA Deworm3 Precision Simulation"
author: "Calum Davey"
date: "04/04/2019"
output: 
  html_document:
    code_folding: hide
---
## Set the design parameters for the study 

The first stage is to set the number of disabled and non-disabled children that will be recruited overall, and the number of clusters. 

```{r}
library(openxlsx)
# Load the number of people enumerated in each cluster 
N <- openxlsx::read.xlsx('Cluster Population CE1.xlsx')
N[,3] <- round(N[,2]*(2000/sum(N[,2])),0) # back-calc for the proportion of CWD in each cluster 
N[,4] <- round(N[,2]*(2000/(sum(N[,2]*0.05))),0) # back-calc for the total children in each cluster 
colnames(N) <- c('cluster','N','N_cwd','N_c')
head(N)
# n <- N[,3] # total number of disabled children 
# 
# d <- matrix(,nrow=n*2, ncol=4) # Create data 
# d[,1] <- rep(c(0,1),times=c(n,n)) # Code disabled, and non-disabled (1,0)
# c <- 40 # total number of clusters 
# d[,2] <- rep(rep(seq(1,c,1), times=rep(n/c, times=c)),times=2)
```

## Set assummed population-level parameters 

Next, we set the population-level parameters, e.g. the 'true' paramter values, for each cluster. In this case, that means the true proportion attending school for disabled and non-disabled children. 

```{r}
# Enrolment 
  e0 <- c(.7) # enrolment of all children 
  e1 <- c(.5) # enrolment of disabled children 
  k <- c(.01) # Set intercluster coefficient of variation in enrollment levels 
  var_b1 <- k/e1 # Calculate the variance of the cluster proportions 
  var_b2 <- k/e0 # Calculate the variance of the cluster proportions 
  # Fix true cluster-level means of school enrollment 
  N$e_1 <- round(rnorm(nrow(N), e1, var_b1^.5),2) # Do need a way to avoid values >1 or <0
  N$e_0 <- round(rnorm(nrow(N), e0, var_b2^.5),2) # Do need a way to avoid values >1 or <0
  # d[,3] <- c(rep(e0_c, times=rep(n/c, times=c)),
  #            rep(e1_c, times=rep(n/c, times=c)))
# True relationship between enrollment and treatment 
  t0 <- c(.8) # 80% of enrolled children are in school and treated 
  t1 <- c(.6) # 60% of enrolled children with disabilities are in school and treated 
# True proportions treated by the intervention
  
```

## Generate enrolment outcome 

Finally, we take 1000 samples from the population, essentially simulating doing the planned study 1000 times. 

```{r}
i <- 1
while (i<=1000){
  d <- cbind(d, apply(d, 1, function(x) sample(c(0,1),size=1,prob=c(1-x[3],x[3]))))
  i <- i+1
}
```

## Conduct analysis on simulated data 

Using the 1000 simulations, we can conduct whatever analysis we want and see how the estiamtes vary from one simulated sample to the next. For example, we can estimate the proportion attending and the difference between the proportions for disabled and non disabled children. 

```{r}
# Calculate sample enrolment figures 
p0 <- apply(d[d[,1]==0, 5:ncol(d)], 2, mean)
p1 <- apply(d[d[,1]==1, 5:ncol(d)], 2, mean)
diff <- apply(cbind(p0,p1),1,function(x) x[1]-x[2])
```

We can visualise the level of precision by plotting the results from all 1000 samples. Here we have the proportion in school for children with disabilities, children without disabilities (controls), and the difference between these for each run of the simulation. 

```{r}
hist(p0, xlim=c(0.1,.8),ylim=c(0,50), main='',
     xlab='R or RD', freq=F)
hist(p1, add =T, freq=F)
hist(diff, xlab = 'RD',add=T, freq=F)
lines(density(diff),col='red')
text(x=c(mean(diff),mean(p1),mean(p0)), 
     y=c(50),
     labels=c('Prop.-diff.', 'Prop. CWD',
              'Prop. controls'))
```





