---
title: "Part 2 - responses"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(global.device = TRUE)
```

In the [previous post](https://calumdavey.github.io/Covid-Deaths/) I showed how to use [data from Johns Hokins](https://github.com/CSSEGISandData/COVID-19) to create a graph of the deaths from Covid-19.

In this post, I'll show how we can use data from UNESCO to mark the dates when schools closed. 

First, we need to run the [previous .R script](https://github.com/calumdavey/calumdavey.github.io/blob/master/files/COVID19/COVID_DEATHS_GRAPHS.R) and use `read.csv()` to load the data. 

```{r Loaddata}
# Run the previous .R script 
source('COVID_DEATHS_GRAPHS.R')
# Load data 
data_raw <- read.csv('https://en.unesco.org/sites/default/files/covid_impact_education.csv', as.is = TRUE)
```

As before, there are various data-management steps.
This time we need to change the names of the UK and the US to match the previous dataset, and only keep the countries we want to use. 

```{r data-management}
# Re-name UK and US 
data_raw$Country[data_raw$Country=='United Kingdom of Great Britain and Northern Ireland'] <- 'United Kingdom'
data_raw$Country[data_raw$Country=='United States of America'] <- 'US'

# Keep only the selected countries 
schools <- data_raw[data_raw$Country %in% countries &
                    data_raw$Scale == 'National', ]
```

We can now add the data to the plot from before. 
The code from the previous plot is repeated below, with an additional few steps in the `for()` loop that produced the trend lines.

```{r Addtoplot}
# Start with an 'empty' plot of Italy 
plot_data <- data[data$Group.1=='Italy', ]

plot(plot_data$x, type = 'n', 
     log = 'y', # y-axis is on the log scale 
     bty = 'n', # no border around the plot 
     xlim = c(1,(nrow(plot_data)+4)),
     axes = FALSE, bg='gray80',
     xlab='Days after 10 confirmed deaths',
     ylab='Confirmed deaths (log scale)',
     sub = 'N.B. US has not closed schools; Japan closed before 10 deaths',
     cex.sub=.6,
     cex.lab=.7)

# Add gridlines 
grid(nx = NULL, ny = NULL, col = "lightgray", lty = "dotted", lwd = par("lwd"), equilogs = F)

# Add each of the other countries 
for (country in countries){
  plot_data <- data[data$Group.1 == country,]
  # Plot the lines 
  lines(c(1:nrow(plot_data)), plot_data$x, 
        col=cols[which(countries==country)])
  # Add the School closures
  # Skip the US, because schools haven't closed 
  if (country != 'US'){
    # Create marker for when schools first closed
    closed <- schools[schools$Country==country,]
    closed <- data.frame(date = as.Date(closed[1,1],
                         '%d/%m/%Y'), now=1)
    
    # Merge the marker with the plot data 
    plot_data_s <- merge(plot_data, closed,
                         by.x='Group.2',
                         by.y='date',
                         all.x=T)
    
    # Identify day and deaths when schools closed
    closed <- c(which(plot_data_s$now==1),
          plot_data_s$x[which(plot_data_s$now==1)])
    
    # Add the lines 
    lines(x=c(closed[1],closed[1]),
          y=c(5, closed[2]),
          col=cols[which(countries==country)],
          lty=2, lwd=.7)
    lines(x=c(0,closed[1]),
          y=c(closed[2], closed[2]),
          col=cols[which(countries==country)],
          lty=2, lwd=.7)
  }
  # Add a label
  text(x=nrow(plot_data), y=max(plot_data$x),
       label=paste0(country,' (',nrow(plot_data),' days)'), 
       pos=4, offset=.1, cex=.6, font=2,
       col=cols[which(countries==country)])
}

# Add the axes 
axis(1, lwd=0, cex.axis=.7)
axis(2, lwd=0, las=1, cex.axis=.7)

# Add titles
mytitle = "Covid-19 deaths"
mysubtitle = "Arranged by number of days since 10 or more deaths"
mtext(side=3, line=2, at=-0.07, adj=0, cex=1, mytitle)
mtext(side=3, line=1, at=-0.07, adj=0, cex=0.7, mysubtitle)

# Add a legend 
legend(x='bottomright', 
       legend=c('Schools closed'), lty=2,
       bty='n', cex=.7)
```

